The function--sheaf dictionary for smooth commutative group schemes over finite fields, with applications over local fields

### 2014-08-21 dictionary.tex
Making notes for the new introduction.

We close the introduction with an overview of the structure of this paper.
After setting some basic notation in Section~\ref{sec:notation},
in Section~\ref{sec:category} we define the category of quasicharacter sheaves.
We define pullbacks and products of quasicharacter sheaves in Section~\ref{sec:pullback}
Section~\ref{sec:Frob} introduces Trace of Frobenius,
one direction of the function--sheaf dictionary.
, and use
them to prove a functoriality result on Trace of Frobenius.
In Section~\ref{sec:basechange} we study how quasicharacter sheaves behave under base change
and Weil restriction.
Sections~\ref{sec:bounded} and \ref{sec:finite} describe bounded quasicharacter sheaves and finite quasicharacter sheaves, respectively,
forming two important subcategories of $\QC(G)$.
%These subcategories play a key role in the proof of Proposition~\ref{prop:restriction}.
In Section \ref{sec:connected} we show that all quasicharacter sheaves on $G$ are finite quasicharacter sheaves when $G$ is connected, and that Trace of Frobenius is an isomorphism in this case.
We shift to \'etale group schemes $G$ in Section~\ref{sec:etale} and find a cohomological interpretation
of quasicharacter sheaves in this case.
In Section~\ref{sec:restriction} we show that for non-connected $G$, restriction to the identity component
induces an essentially surjective functor of quasicharacter sheaves.
%Section~\ref{sec:snake} presents the main theorem of this paper: Trace of Frobenius
%is a functorial isomorphism from isomorphism classes of flabby quasicharacter sheaves on $G$
%to characters of $G(\Fq)$.

@@

\section{Quasicharacter sheaves on connected commutative algebraic groups}

Before turning our attention to quasicharacter sheaves on connected commutative algebraic groups in Section~\ref{sec:connected}, we being with two sections that treat two special classes of quasicharacter sheaves on arbitrary smooth commutative group schemes $G$: those that descend from local systems on $\bG$ to local systems on $G$ in Section~\ref{sec:bounded}; and those local systems on $G$ that are defined by discrete isogenies onto $G$ in Section~\ref{sec:discrete_isogenies}. The latter will play a role in Sections~\ref{sec:connected} and \ref{sec:restriction}.

\subsection{Descent to G}\label{sec:bounded}

@@

  Since $b_G^*$ is full and faithful
  \citelist{\cite{deligne-katz:SGA7.2}*{Expos\'e XIII} \cite{beilinson-bernstein-deligne:81a}*{Prop. 5.2.1}},
  $
  b_{G\times G}^* : \Hom(m^*\qcs{E},\qcs{E}\boxtimes\qcs{E}) \to \Hom({\bar m}^*\gqcs{L},\gqcs{L}\boxtimes\gqcs{L})
  $
  is a bijection
  (hom taken in the categories on constructible $\ell$-adic sheaves on
  $G\times G$ and ${\bar G}\times {\bar G}$ respectively,
  in which $\ell$-adic local systems sit as full subcategories).
  Let $\mu_\qcs{E} : m^*\qcs{E} \to \qcs{E}\boxtimes\qcs{E}$ be the isomorphism matching
  $\mu : {\bar m}^*\gqcs{L} \to \gqcs{L}\boxtimes\gqcs{L}$,
  the latter appearing in the definition of $\qcs{L}$.
  Then, as in Section~\ref{sec:bounded}, $(\qcs{E},\mu_\qcs{E})$ is an object in $\bQC(G)$
  and $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ is isomorphic to $(b_G^*\qcs{E},b_{G\times G}^*\mu_\qcs{E})$ in $\QC(G)$.
  Thus, the full and faithful functor $B_G : \bQC(G) \to \QC(G)$ from Section~\ref{sec:bounded}
  is also essentially surjective, hence an equivalence.
\end{proof}

@@

>>>>

We close the introduction with an overview of the structure of this paper.
After setting some basic notation in Section~\ref{sec:notation},
in Section~\ref{sec:category} we define the category of quasicharacter sheaves.
We define pullbacks and products of quasicharacter sheaves in Section~\ref{sec:pullback}
Section~\ref{sec:Frob} introduces Trace of Frobenius,
one direction of the function--sheaf dictionary.
, and use
them to prove a functoriality result on Trace of Frobenius.
In Section~\ref{sec:basechange} we study how quasicharacter sheaves behave under base change
and Weil restriction.
Sections~\ref{sec:descentG} and \ref{sec:finite} describe bounded quasicharacter sheaves and finite quasicharacter sheaves, respectively,
forming two important subcategories of $\QC(G)$.
%These subcategories play a key role in the proof of Proposition~\ref{prop:restriction}.
In Section \ref{sec:connected} we show that all quasicharacter sheaves on $G$ are finite quasicharacter sheaves when $G$ is connected, and that Trace of Frobenius is an isomorphism in this case.
We shift to \'etale group schemes $G$ in Section~\ref{sec:etale} and find a cohomological interpretation
of quasicharacter sheaves in this case.
In Section~\ref{sec:restriction} we show that for non-connected $G$, restriction to the identity component
induces an essentially surjective functor of quasicharacter sheaves.
%Section~\ref{sec:snake} presents the main theorem of this paper: Trace of Frobenius
%is a functorial isomorphism from isomorphism classes of flabby quasicharacter sheaves on $G$
%to characters of $G(\Fq)$.

@@

\section{Quasicharacter sheaves defined by discrete isogenies}

Before turning our attention to quasicharacter sheaves on connected commutative algebraic groups in Section~\ref{sec:connected}, we treat two special classes of quasicharacter sheaves on arbitrary smooth commutative group schemes $G$:  in Section~\ref{sec:descentG}, those that descend from local systems on $\bG$ to local systems on $G$; and in Section~\ref{sec:discrete_isogenies}, those that are defined by discrete isogenies onto $G$. The latter will play a role in Sections~\ref{sec:connected} and \ref{sec:restriction}.

\subsection{Descent to G}\label{sec:descentG}

@@

  Since $b_G^*$ is full and faithful
  \citelist{\cite{deligne-katz:SGA7.2}*{Expos\'e XIII} \cite{beilinson-bernstein-deligne:81a}*{Prop. 5.2.1}},
  $
  b_{G\times G}^* : \Hom(m^*\qcs{E},\qcs{E}\boxtimes\qcs{E}) \to \Hom({\bar m}^*\gqcs{L},\gqcs{L}\boxtimes\gqcs{L})
  $
  is a bijection
  (hom taken in the categories on constructible $\ell$-adic sheaves on
  $G\times G$ and ${\bar G}\times {\bar G}$ respectively,
  in which $\ell$-adic local systems sit as full subcategories).
  Let $\mu_\qcs{E} : m^*\qcs{E} \to \qcs{E}\boxtimes\qcs{E}$ be the isomorphism matching
  $\mu : {\bar m}^*\gqcs{L} \to \gqcs{L}\boxtimes\gqcs{L}$,
  the latter appearing in the definition of $\qcs{L}$.
  Then, as in Section~\ref{sec:descentG}, $(\qcs{E},\mu_\qcs{E})$ is an object in $\bQC(G)$
  and $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ is isomorphic to $(b_G^*\qcs{E},b_{G\times G}^*\mu_\qcs{E})$ in $\QC(G)$.
  Thus, the full and faithful functor $B_G : \bQC(G) \to \QC(G)$ from Section~\ref{sec:descentG}
  is also essentially surjective, hence an equivalence.
\end{proof}

@@

\section{Notes for the new Introduction}

Motivated to an application to the geometrization and categorification of quasicharacters of p-adic tori,
this paper extends Deligne's function--sheaf dictionary for connected commutative algebraic groups over finite fields to all smooth commutative group schemes over finite fields.

As Deligne explained in \cite{deligne:SGA4.5}*{sommes trig.}, if $G$ is a connected commutative algebraic group over $\FF_q$, then the trace of Frobenius provides a perfect dictionary between characters of $G(\FF_q)$ and isomorphism classes of rank-one $\ell$-adic local systems $\mathcal{E}$ on $G$ for which $m^*\qcs{E} \iso \qcs{E} \boxtimes \qcs{E}$, where $m : G\times G\to G$ is the multiplication map.
If one wishes to make a category from this class of local systems, one is lead to consider sheaf morphisms $\qcs{E} \to \qcs{E}'$ which are compatible with particular choices of isomorphisms $m^*\qcs{E} \iso \qcs{E} \boxtimes \qcs{E}$ and $m^*\qcs{E}' \iso \qcs{E}'\boxtimes \qcs{E}'$. {\it A priori}, these sheaf morphisms do not form a category, since
